---
title: "Lecture Prep 07 Replication"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This is the replication notebook for LP-07 from the course SOC 4650/5650: Introduction to GISc. 

## Load Dependencies
The following code loads the package dependencies for our analysis:

```{r package-load}
# tidyverse
library(readr) # reading/writing csv files

# other packages
library(here) # working directory tools
library(sf) # spatial data tools
```

## Load Data
The following code loads the data package and assigns our data to a data frame in our global environment. The `stl_as_sf()` function can be used to convert certain tables in stlData into `sf` objects.

```{r data-load}
library(stlData)
leadTibble <- stl_tbl_lead
leadSf <- stl_as_sf(stl_tbl_lead)
```

## Part 1
### Question 2
The `write_csv()` function from `readr` can be used to write tabular data from `R`:

```{r}
write_csv(leadTibble, here("data", "leadTibble.csv"))
```

We now have a `.csv` file saved in our data folder containing the `leadTibble` data.

## Part 2
### Question 4
The `st_write()` function from `sf` can be used to write shapefile data from `R`:

```{r}
st_write(leadSf, here("data", "leadData.shp"), delete_dsn = TRUE)
```

We now have a `.shp` file saved in our data folder containing the `leadSf` data. If you try and knit your notebook without the `delete_dsn` option, you will get an error that the data already exist and the file will not be created. If you try and execute your code with the `delete_dsn` option *and* you have not already done so, you will get an error that the file does not appear in the specified directory. However, the file *will* still be created.
